A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-03 shanghai.tex
Edits, to section 8

\begin{lemma}\label{lemma:ext}
Every discrete isogeny onto $G^0$ extends to a discrete
isogeny onto $G$ such that the component group of the covering space is isomorphic
to the component group of $G$.
\end{lemma}

\begin{proof}
Let $\pi: B \to G^0$ be a discrete isogeny; set $A \ceq \ker \pi$.
  We will find a discrete isogeny $f: H\to G$
  such that that $H^0 = B$, $f^0 =\pi$ and
  $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism of component
  groups.
  \begin{equation}\label{extension-diagram}
  \begin{tikzcd}
  A \arrow[equal]{r} \dar & A \dar \\
  B \rar \dar[swap]{\pi} & H \rar \dar[swap]{f} & \pi_0(H) \arrow{d}[below,rotate=90]{\sim}[swap]{\pi_0(f)} \\
  G^0 \rar & G \rar & \pi_0(G)
  \end{tikzcd}
  \end{equation}
  where all rows and columns are exact and all maps are defined over
  $\Fq$.  We will do so by passing back and forth between group
  schemes over $\Fq$ and their $\bFq$-points.

  Set $\Gamma \ceq \Gal(\bFq/\Fq)$. Extensions of
  $G^0(\bFq)$ by $A(\bFq)$, such as $A(\bFq) \to H^0(\bFq) \to G^0(\bFq)$,
  correspond to classes in $\Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$.
  Similarly, extensions of $G(\bFq)$ by $A(\bFq)$ correspond to
  classes in $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq))$.  The map
  $G^0(\bFq) \to G(\bFq)$ induces the map
  \[
  \Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))
  \]
  fitting into the long exact sequence derived from applying
  the functor $\Hom(\mbox{---}, A(\bFq))$ to $G^0(\bFq) \to G(\bFq) \to \pi_0(G)(\bFq)$:
  \[
  \Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq)) \to \Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)).
  \]

>>>>

\begin{lemma}\label{lemma:ext}
Every discrete isogeny onto $G^0$ extends to a discrete
isogeny onto $G$ inducing an isomorphism on component groups.
\end{lemma}

\begin{proof}
Let $\pi: B \to G^0$ be a discrete isogeny; set $A \ceq \ker \pi$.
  We will find a discrete isogeny $f: H\to G$
  such that that $H^0 = B$, $f^0 =\pi$ and
  $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism of component
  groups.  Namely, we will fit $\pi$ into the following diagram,
  \begin{equation}\label{extension-diagram}
  \begin{tikzcd}
  A \arrow[equal]{r} \dar & A \dar \\
  B \rar \dar[swap]{\pi} & H \rar \dar[swap]{f} & \pi_0(H) \arrow{d}[below,rotate=90]{\sim}[swap]{\pi_0(f)} \\
  G^0 \rar & G \rar & \pi_0(G),
  \end{tikzcd}
  \end{equation}
  where all rows and columns are exact and all maps are defined over
  $\Fq$.  We will do so by passing back and forth between group
  schemes over $\Fq$ and their $\bFq$-points.

  Set $\Gamma \ceq \Gal(\bFq/\Fq)$. Extensions of
  $G^0(\bFq)$ by $A(\bFq)$, such as $A(\bFq) \to B(\bFq) \to G^0(\bFq)$,
  correspond to classes in $\Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$.
  Similarly, extensions of $G(\bFq)$ by $A(\bFq)$ correspond to
  classes in $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq))$.  The map
  $G^0(\bFq) \to G(\bFq)$ induces the map
  \[
  \Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))
  \]
  fitting into the long exact sequence derived from applying
  the functor $\Hom(\mbox{---}, A(\bFq))$ to $G^0(\bFq) \to G(\bFq) \to \pi_0(G)(\bFq)$:
  \[
  \Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq)) \to \Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)).
  \]
